Report on Late Winter2013 Production: Image Differencing

### 2013-03-29 W13report.tex
Added discussion of fp vs threshold and some discussion of the MSE work

\usepackage[paperwidth=8.5in,paperheight=11in,centering,margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amsbsy}
\usepackage{natbib}
\usepackage{rotating}
\input epsf
\usepackage{amsmath}
\usepackage{wasysym}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{color}
%\usepackage{ulem}
%\usepackage{epstopdf}
\usepackage[toc,page]{appendix}

@@

We measured flux of {\tt DiaSources} on postfiltered difference images
using PSF--weighted flux photometry (as implemented in the LSST
measurement algorithm "flux.psf").  For measuring fluxes of point
sources on prefiltered images we implemented a new algorithm
"flux.filtered", which works as follows:
%
\[flux = \frac{\sum_{(x,y)}I(x,y) * \phi(x,y)}{\sum_{(x',y')}\phi^2(x',y')}\]
%

where $I(x,y)$ is the image before PSF filtering and $\phi(x,y)$ is an
image of the PSF centered on the source. This equation assumes an
isolated point source, an assumption that is reasonable for difference
images.  We do not weight by the per--pixel variance to avoid
non--linearity in the flux measurement, at the expense of less precise
measurements.

Filtering is performed by convolving the image with its PSF, so the
peak of the filtered image of a given source is:
%
\[Ipeak_{filt} = \sum_{(x,y)}I(x,y) * \phi(x,y)\]
%
Hence the equation for measuring its flux reduces to:
%
\[flux = \frac{Ipeak_{filt}}{\sum_{(x',y')}\phi^2(x',y')}\]
%
Similarly, the predicted uncertainty of the flux measurement is:
%
\[\sigma = \frac{\sqrt{\sum_{x,y}\sigma{^2(x,y)}\phi^2(x,y)}}{\sum_{x',y'}\phi^2(x',y')}\]
%
where $\sigma{^2(x,y)}$ is the is the per-–pixel variance of the
unfiltered image. This reduces to the following:
%
\[\sigma = \frac{\sqrt{\sigma peak{_{filt}^2}}}{\sum_{x',y'}\phi^2(x',y')}\]
%
where $\sigma peak{_{filt}^2}$ is the value of the variance plane,
after filtering, at the peak of the source.

The center of each source is determined using an LSST adaptation of
the SDSS centroiding algorithm "centroid.sdss". In general the source
will not be centered on a pixel, so "flux.filtered" uses a Lanczos 4
resampling kernel to shift the source to be centered on the nearest
pixel before applying the equations above.

@@

\subsection{Effect of Uncertainty on Measured ${S}\over{N}$ {\bf SIMON}}

@@

\subsection{Correlation with KernelCandidate Statistics {\bf ANDYB}}

@@

@@

\item Our empirical negative to positive false detection ratio is {\bf
  X.X}, indicating a bias in the background levels of {\bf X\%}.

@@

\section{Outstanding Issues {\bf SIMON}}
We ran into several issues during this production that required work--arounds.  
In some cases we did not have the time to drill all the way down
to the root cause.  We itemize these issues here so that they may be addressed 
as part of the Summer 2013 work.

\subsection{Astrometric Registration \label{subsec-astrom}}
The original plan for this production was to use the single deep exposures (v6866600) as 
inputs to the coadd process, to exercise use of the coadd repository.  As 
such we warped the exposures to coadd patches for use in the image difference
task.  The register task was unable to solve a differential WCS between the coadd
and science {\tt calexp} because in coadds, the parent WCS is for the entire {\tt tract}.
Each coadd {\tt patch} thus has a CRPIX at the center of the {\tt tract}, which is
typically off the {\tt patch} image.  TAN--SIP is unable to perform in this circumstance,
as it expands its polynomials around the CRPIX.  To get around this 
we tried re--solving the coadd patches using a TAN--SIP solution that was solved from
scratch.  This worked fairly will part of the time,
yielding residuals in the 40--100 mas range, compared to the 250 mas previously.  However,
as we did not understand the origin of these problems, and there were legitimate bugs
in meas\_astrom  that were uncovered as a result of this investigation, we reverted to
using the deep {\tt calexp} as the template directly; the direct {\tt calexp} to {\tt calexp} 
registration yielded residuals in the range of 5--7 mas (well, its 0.02 pixels).

@@

\begin{table}
\centering
\begin{tabular}{lcc}
\hline
\multicolumn{3}{|c|}{Theoretical Number of False Detections per Sensor} \\
\hline
Visit    & FWHM (pixels) & N\\
\hline
v6866601 & 3.0 & N \\
v6866602 & 4.4 & N \\
v6866603 & 6.0 & N \\
\end{tabular}
\caption{Total number of false detections that we expect based on
  {\bf blah}.  We list the total number of positive {\it plus} negative detections;
  they should be present in equal quantities. \label{tab-fp}
}
\end{table}

@@

>>>>

\usepackage[paperwidth=8.5in,paperheight=11in,centering,margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amsbsy}
\usepackage{natbib}
\usepackage{rotating}
\input epsf
\usepackage{amsmath}
\usepackage{caption}
\usepackage{wasysym}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{color}
%\usepackage{ulem}
%\usepackage{epstopdf}
\usepackage[toc,page]{appendix}

@@

Compare the measured S/N of the DiaSources with the detection threshold.

@@

\subsection{Effect of Uncertainty on Measured ${S}\over{N}$ {\bf SIMON}\label{sec-sn}}

@@

\subsection{Detection Threshold}
Obviously the number of false positives is strongly dependent on the detection threshold.  In order to
check the scaling of our false positive rates with detection threshold, we ran the same config (see \ref{tab-bestconfig10} 
for the config parameters)
at several detection threshold values for each visit both prefiltering and postfiltering.  As discussed in 
Section~\ref{sec-analyticfp} one can predict the number of false positives expected in a sensor given
seeing and threshold values.  Some example values of the number of false positives (both +ve and -ve)
expected in LSST size detectors are given in Table \ref{tab-fprate}.

As discussed in the Section \ref{sec-sn}, the accuracy of the noise estimation in the dection is crutial to 
understanding the false positive rate.  In the previous section we quantify the under--estimation of the variance
by the variance plane in the case of pre and post filtering.  In our analysis of the threshold dependence of
our false detection rate we must correct for this under--estimation in order to put the counts from each algorithm
on the same normalization.

Figure \ref{fig-fp_v_thresh} shows the results of this analysis.  In all figures red squares are postfiltered data and
blue circles are for the prefiltering technique.  All values are for the number of false positives in a 4000x4072 sensor.
Using a Monte Carlo approach we 
estimate the variance on the predicted false positive rate as a function of threshold.  The 1-$\sigma$ confidence envelope is shown
as shaded grey about the model line in dark black.  From left to right the panels are for visit 6866601 -- 0.6$^{\prime\prime}$ FWHM,
6866602 -- 0.88$^{\prime\prime}$ FWHM and 6866603 -- 1.2$^{\prime\prime}$ FWHM respectively.  Except in the case of deconovlution
in the postfiltering case (visit 6866601) the count rates between the two algorithms agree very well when corrected for the
observed under--estimation of the noise by the variance plane.  In the case where postfiltering requires a deconvolution, the high $\over{S}{N}$
detections come primarily from detection of the ringing around stars.

In all cases the empirical results are below the predicted results indicating
that the number counts are smaller than the theoretical floor.  {\bf What should we say about this here?} We can correct for this by scaling
the prediction by 2.5\%, 1.5\% and 1\% for visits 6866601, 6866602, and 6866603 respectively.  The data plotted with these corrections are shown
in the bottom row.  Once this correction is done, the trend of false positive rate with detection threshold agrees to within 1-$\sigma$ for the 
prefiltering case for all three seeing values.

\section{Results: Correlation with KernelCandidate Statistics {\bf ANDYB}}

@@

Using this simplified data set, we can begin to look at how we may predict 
optimal configs based measured statistics.  We notice that the MSE of the
control sample has a minimum in all cases.  See Figures \ref{fig-MSE_CONTROL}
and \ref{fig-MSE_FITTED} to see the behavior of the MSE for the control sample and
fitted sample respectively.  We note that the config at the minimum MSE value is
not always the same config.  Table \ref{tab-bestconfig_MSE} lists the configs
at the minimum of the MSE for the control sample.  This can be compared to
the configs in Table \ref{tab-bestconfig10}.

Since we were looking as MSE values for all chips at once, it's possible that
each sensor has it's own optimal config and thus the aggregate MSE is not a 
good indicator of overall config optimization.  To get a handle on this, we
looked at per sensor MSE values and per sensor false positive rates.  In fact,
the config that produces the minimum number of false positives is variable
from sensor to sensor.  It is worth investigating how much the false positive rate
per chip goes down if per sensor optimal configs are used instead of per visit optimal
configs.

When we look at the MSE per chip and compare with the minimum false positive config we 
find good agreement.  The findings are summarized in Table \ref{tab-MSE_persensor}.  Because
the MSE is sensitive to over fitting, we observe that it tends to lower complexity configs
even if the false positive rate is slightly higher.

@@

and \ref{fig-MSE_FITTED} to see the behavior of the MSE for the control sample and
fitted sample respectively.  We note that the config at the minimum MSE value is
not always the same config\item Our empirical negative to positive false detection ratio is {\bf
  X.X}, indicating a bias in the background levels of {\bf X\%}.

@@

\section{Outstanding Issues {\bf SIMON}}
We ran into some issues during this production that required work--arounds.  
In some cases we did not have the time to drill all the way down
to the root cause.  We itemize these issues here so that they may be addressed 
as part of the Summer 2013 work.

\subsection{Astrometric Registration \label{subsec-astrom}}
The original plan for this production was to use the single deep exposures (v6866600) as 
inputs to the coadd process, to exercise use of the coadd repository.  As 
such we warped the exposures to coadd patches for use in the image difference
task.  The register task was unable to solve a differential WCS between the coadd
and science {\tt calexp} because in coadds, the parent WCS is for the entire {\tt tract}.
Each coadd {\tt patch} thus has a CRPIX at the center of the {\tt tract}, which is
typically off the {\tt patch} image.  TAN--SIP is unable to perform in this circumstance,
as it expands its polynomials around the CRPIX.  To get around this 
we tried re--solving the coadd patches using a TAN--SIP solution that was solved from
scratch.  This worked fairly will part of the time,
yielding residuals in the 40--100 mas range, compared to the 250 mas previously.  However,
as we did not understand the origin of these problems, and there were legitimate bugs
in meas\_astrom  that were uncovered as a result of this investigation, we reverted to
using the deep {\tt calexp} as the template directly; the direct deep {\tt calexp} to science {\tt calexp} 
registration typically yielded residuals in the range of 4--5 mas.

@@

\begin{table}
\centering
\begin{tabular}{lcc}
\hline
\multicolumn{3}{|c|}{Theoretical Number of False Detections per Sensor} \\
\hline
Visit    & FWHM (pixels) & N\\
\hline
v6866601 & 3.0 & N \\
v6866602 & 4.4 & N \\
v6866603 & 6.0 & N \\
\end{tabular}
\caption{Total number of false detections that we expect based on
  {\bf blah}.  We list the total number of positve {\it plus} negative detections;
  they should be present in equal quantities. \label{tab-fp}
}
\end{table}

@@

\begin{figure}%
\captionsetup[subfloat]{labelformat=empty}
\centering
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_01_nocorr.eps}}
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_02_nocorr.eps}}
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_03_nocorr.eps}}\\
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_01_corr.eps}}%
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_02_corr.eps}}%
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_03_corr.eps}}%
\caption{}
\label{fig-fp_v_thresh}
\end{figure}
